Presolving for\\ Mixed-Integer Semidefinite Optimization

### 2020-12-10 tex/MISDP-Presolving/MISDPpresolving.tex
add some remarks to presolving document

The following inequalities are known from the literature, see
Mars~\cite{Mar13} and  Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
\eqref{MISDP}:
\begin{itemize}
\item Any feasible $y \in \R^m$ satisfies $A(y) \succeq 0$, which implies
  that the diagonal entries of $A(y)$ are nonnegative, i.e., for all
  $i \in [n]$:
  \begin{align}\label{eq:diaggezero}
    \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
  \end{align}
\item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
  $A^k_{ij} = 0$ for all $k \in [m]$, $A^k_{ii} = 0$ for all continuous
  variables and $\ell_i \geq 0$ for all integer variables,
  the following inequality is valid:
  \begin{align}\label{eq:diagzeroimpl}
    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq1.
  \end{align}
  Note that the conditions imply that $A^0$ is not psd.
\item For all $X \succeq 0$, we have
  \begin{align}\label{eq:2minorlinP}
    X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
    X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
  \end{align}
  This follows by using a 2 by 2 minor and multiplying from left
  and right by the all-ones vector and $[1,-1]$, respectively. It also
  follows from~\eqref{eq:TwoByTwoMinorInequality}.
\item The previous inequalities translated to the matrix pencil are:
  \begin{equation}\label{eq:2minorlinD}
    \begin{aligned}
      & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
      \quad\Leftrightarrow\quad& \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} -
      2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
    \end{aligned}
  \end{equation}
\item If $A^k_{ii} = A^k_{jj} = 0$ for all~$k \in [m]$
  and~$A^0_{ii}A^0_{jj} \geq 0$ for some $i \neq j \in [n]$, then the
  following inequality is valid:
  \begin{align}\label{eq:2minorprodD}
    \sum_{k=1}^m A^k_{ij} y_k \geq A^0 - \sqrt{A^0_{ii}A^0_{jj}}.
  \end{align}
  Furhtermore, if exactly one~$A^k_{ij} \neq 0$, then this yields upper or
  lower bounds for the corresponding variable~$y_k$, depending on the sign
  of~$A^k_{ij}$.
\item For a positive semidefinite matrix $X$ the following quadratic
  inequality holds by using the 2 by 2 minor:
  $X_{st}^2 \leq X_{ss} X_{tt}$. This constraint is SOC representable:
  \begin{align*}
    \left\|\binom{2\,X_{st}}{X_{ss} - X_{tt}}\right\| &\leq X_{ss} +
                                                        X_{tt}. \\
    \quad\Leftrightarrow\quad 4\, X_{st}^2 + (X_{ss}^2 - 2\,X_{ss}\, X_{tt}
    + X_{tt}^2) & \leq X_{ss}^2 + 2\, X_{ss}\,X_{tt} + X_{tt}^2.
  \end{align*}
  Translated to the matrix pencil notation the cut looks as follows:
  \begin{equation}\label{eq:2minorSOC}
    \begin{aligned}
      &\left\|\binom{2\, \sum_{i=1}^m (A_i)_{st}\, y_i - 2\, (A_0)_{st}}{ q
          \sum_{i=1}^m (A_i)_{ss}\, y_i - (A_0)_{ss} - \sum_{i=1}^m
          (A_i)_{tt}\, y_i + (A_0)_{tt}}\right\| \\ &\leq \sum_{i=1}^m
      (A_i)_{ss}\, y_i - (A_0)_{ss} + \sum_{i=1}^m (A_i)_{tt}\, y_i -
      (A_0)_{tt}.
    \end{aligned}
  \end{equation}
  These constraints can be added directly as SOC-constraints.
\end{itemize}

@@

@@

  Upper bounding via
  McCormick inequalities yields:
  \begin{align*}
    X_{ii} X_{jj} & \leq \min \{L_{jj}\, X_{ii} + U_{ii}\, X_{jj} - U_{ii}\, L_{jj},\;
                    U_{jj}\, X_{ii} + L_{ii}\, X_{jj} - L_{ii}\, U_{jj}\}  \\
    & \leq \min \{L_{jj}\, U_{ii} + U_{ii}\, U_{jj} - U_{ii}\, L_{jj},\;
                    U_{jj}\, U_{ii} + L_{ii}\, U_{jj} - L_{ii}\, U_{jj}\}  \\
    & = \min \{U_{ii}\, U_{jj},\;  U_{jj}\, U_{ii}\} = U_{ii}\, U_{jj}.
  \end{align*}
  Taking the square root gives the claim.
\end{proof}

@@

>>>>

The following inequalities are known from the literature, see
Mars~\cite{Mar13} and  Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
\eqref{MISDP}:
\begin{itemize}
\item Any feasible $y \in \R^m$ satisfies $A(y) \succeq 0$, which implies
  that the diagonal entries of $A(y)$ are nonnegative, i.e., for all
  $i \in [n]$:
  \begin{align}\label{eq:diaggezero}
    \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
  \end{align}
\item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
  $A^k_{ij} = 0$ for all $k \in [m]$, $A^k_{ii} = 0$ for all continuous
  variables and $\ell_i \geq 0$ for all integer variables,
  the following inequality is valid:
  \begin{align}\label{eq:diagzeroimpl}
    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq1.
  \end{align}
  Note that the conditions imply that $A^0$ is not psd.
\item For all $X \succeq 0$, we have
  \begin{align}\label{eq:2minorlinP}
    X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
    X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
  \end{align}
  This follows by using a 2 by 2 minor and multiplying from left
  and right by the all-ones vector and $[1,-1]$, respectively. It also
  follows from~\eqref{eq:TwoByTwoMinorInequality}.
  Note that an additional rank-1 constraint on the matrix~$X$ does not lead
  to stronger inequalities.
\item The previous inequalities translated to the matrix pencil are:
  \begin{equation}\label{eq:2minorlinD}
    \begin{aligned}
      & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
      \quad\Leftrightarrow\quad& \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} -
      2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
    \end{aligned}
  \end{equation}
\item If $A^k_{ii} = A^k_{jj} = 0$ for all~$k \in [m]$
  and~$A^0_{ii}A^0_{jj} \geq 0$ for some $i \neq j \in [n]$, then the
  following inequality is valid:
  \begin{align}\label{eq:2minorprodD}
    \sum_{k=1}^m A^k_{ij} y_k \geq A^0 - \sqrt{A^0_{ii}A^0_{jj}}.
  \end{align}
  Furhtermore, if exactly one~$A^k_{ij} \neq 0$, then this yields upper or
  lower bounds for the corresponding variable~$y_k$, depending on the sign
  of~$A^k_{ij}$.
\item For a positive semidefinite matrix $X$ the following quadratic
  inequality holds by using the 2 by 2 minor:
  $X_{st}^2 \leq X_{ss} X_{tt}$. This constraint is SOC representable:
  \begin{align*}
    \left\|\binom{2\,X_{st}}{X_{ss} - X_{tt}}\right\| &\leq X_{ss} +
                                                        X_{tt}. \\
    \quad\Leftrightarrow\quad 4\, X_{st}^2 + (X_{ss}^2 - 2\,X_{ss}\, X_{tt}
    + X_{tt}^2) & \leq X_{ss}^2 + 2\, X_{ss}\,X_{tt} + X_{tt}^2.
  \end{align*}
  Translated to the matrix pencil notation the cut looks as follows:
  \begin{equation}\label{eq:2minorSOC}
    \begin{aligned}
      &\left\|\binom{2\, \sum_{i=1}^m (A_i)_{st}\, y_i - 2\, (A_0)_{st}}{ q
          \sum_{i=1}^m (A_i)_{ss}\, y_i - (A_0)_{ss} - \sum_{i=1}^m
          (A_i)_{tt}\, y_i + (A_0)_{tt}}\right\| \\ &\leq \sum_{i=1}^m
      (A_i)_{ss}\, y_i - (A_0)_{ss} + \sum_{i=1}^m (A_i)_{tt}\, y_i -
      (A_0)_{tt}.
    \end{aligned}
  \end{equation}
  These constraints can be added directly as SOC-constraints.
\end{itemize}

@@

\begin{remark}
  The assumptions under which inequalities~\eqref{eq:diagzeroimpl} are
  applicable can be weakened. These inequalities are also valid
  if~$A^0_{ii} > 0$, since, under the remaining assumptions,
  \begin{align*}
    0 \leq A(y)_{ii} \cdot A(y)_{jj} - (A^0_{ij})^2 \quad &\implies \quad
    0 < A(y)_{ii} = \sum_{k} A^k_{ii}y_k - A^0_{ii}, \\
    &\implies \quad 0 < \sum_{k} A^k_{ii}y_k
  \end{align*}
  and thus~$0 < \sum_{k} A^k_{ii}y_k$ has to hold. This implies the
  validity of~\eqref{eq:diagzeroimpl}.
\end{remark}

@@

  Upper bounding via
  McCormick inequalities yields:
  \begin{align*}
    X_{ii} X_{jj} & \leq \min \{L_{jj}\, X_{ii} + U_{ii}\, X_{jj} - U_{ii}\, L_{jj},\;
                    U_{jj}\, X_{ii} + L_{ii}\, X_{jj} - L_{ii}\, U_{jj}\}  \\
    & \leq \min \{L_{jj}\, U_{ii} + U_{ii}\, U_{jj} - U_{ii}\, L_{jj},\;
                    U_{jj}\, U_{ii} + L_{ii}\, U_{jj} - L_{ii}\, U_{jj}\}  \\
    & = \min \{U_{ii}\, U_{jj},\;  U_{jj}\, U_{ii}\} = U_{ii}\, U_{jj}.
  \end{align*}
  Taking the square root gives the claim.
\end{proof}
Note that these bounds are tight, even for a rank-1 matrix~$X$: Consider
the all-ones matrix, which clearly has rank~1.

@@

\begin{remark}
  Most likely this can be generalized to the more interesting case of a
  constraint of the form
  \[
  \begin{pmatrix}
    t & x\T \\
    x & X
  \end{pmatrix}
  \succeq 0,
  \quad
  \ell \leq (x,t) \leq u,
\]
where~$t$ is a scalar variable. Since constraints of this particular form
appear in truss topology design and cardinality least-squares problems, we
should check whether this helps for such instances.
\end{remark}
